---
title: "Tplyr table module"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tplyr table module}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This guide provides a detailed overview of the `mod_Tplyr_table` module and its features. It is meant to provide guidance to App Creators on creating Apps in DaVinci using the `mod_Tplyr_table` module.

Walk-throughs for sample app creation using the module are also included
to demonstrate the various module specific features.


`mod_Tplyr_table` serves as a wrapper for displaying interactive summary tables generated with [`{Tplyr}`](https://atorus-research.github.io/Tplyr/). The module leverages Tplyr’s metadata extraction functionality to offer traceability from the summary tables back to the source listing. By clicking on a result cell in the table, the source listing will be generated. For displaying the listing the listings module of  `{dv.listing}` is used, which means all functionality of that module can be leveraged. 

The Tplyr table module allows multiple output tables created with `{Tplyr}` and standalone listings to be displayed under the same tab, using a drop-down selector.

# Features

`mod_Tplyr_table` features the following outputs:

- table created using `Tplyr` package 
- corresponding listing from the selected cell of the table.
- standalone listing

`dv.tables::mod_Tplyr_table()`
module uses several arguments with the following being mandatory and
the rest optional. As part of app creation, the app
creator should specify the values for these arguments as applicable.

**Mandatory Arguments**

-   `module_id` : A unique identifier of type character for the module

- `output_list`: A named list defining the outputs to be generated. Each element of the list corresponds to a table or listing and must be a named list with one of the following structures:

  For summary tables:
    - `tplyr_tab_fun` A function that takes one or more datasets as input and returns a `tplyr_table` object.
    - `build_fun` A function that takes the `tplyr_table` object and returns a built table (typically using `Tplyr::build()`).  
    *Note:  The metadata argument of `Tplyr::build()` needs to be set to `TURE`, so that the corresponding listing can be shown.*  
    
  For standalone listings:
    - `dataset_names` A character vector of dataset names required to generate the listing.
  The names of the top-level list elements are used as identifiers for the outputs.

-  `subjid_var` Column corresponding to subject ID. Default value is 'USUBJID'

**(Optional) Arguments for listing customization**

The module uses `{dv.listings}` for showing drill-down listing data. Therefore additional arguments can be provided for further customization.

`default_vars`, `pagination`, `intended_use_label`, `receiver_id`, `review` are all arguments that are passed through to `{dv.listings}`. more information can be found [here](https://boehringer-ingelheim.github.io/dv.listings/)

# Example

To better understand the set up of the module you can follow the below example:


First load dummy data:

```{r, eval=FALSE}
# load demo data
dm <- pharmaversesdtm::dm
```

Create a Tplyr table and wrap it into a function.

```{r, eval=FALSE}
my_tplyr_fun <- function(dm) {
  tab <- Tplyr::tplyr_table(dm, ARM) |>
    Tplyr::add_layer(Tplyr::group_desc(AGE, by = "Age (years)")) |>
    Tplyr::add_layer(Tplyr::group_count(SEX, by = "Sex")) |>
    Tplyr::add_layer(Tplyr::group_count(RACE, by = "Race"))
  return(tab)
}

```

Create a build function for the tplyr table created in the first step

```{r, eval=FALSE}
build_func <- function(tab) {
  Tplyr::build(tab, metadata = TRUE) |>
    dplyr::mutate(
      row_label2 = ifelse(row_label2 == row_label1, "Total (%)", row_label2)
      ) |>
    Tplyr::apply_row_masks(row_breaks = TRUE)
}
```

Create the output_list passed to `dv.tables::mod_Tplyr_table`
```{r, eval=FALSE}
output_list <- list(
  "Demographic" = list(
    tplyr_tab_fun = my_tplyr_fun,
    build_fun = build_func
  )
)
```

Put everything together and start the app within `{dv.manager}`
```{r, eval=FALSE}

module_list <- list(
  "Table" = dv.tables::mod_Tplyr_table(
    module_id = "tplyr_table",
    output_list = output_list
  )
)

dv.manager::run_app(
  data = list("My data" = list(dm = dm)),
  module_list = module_list,
  filter_data = "dm",
  enable_dataset_filter = TRUE
)

```


It's also possible to add more than one table into the module.

```{r, eval=FALSE}
# load additional demo data
dm <- pharmaversesdtm::dm 

# create second tplyr table
my_tplyr_fun2 <- function(dm, ae) {
  dm_arm <- dm |> dplyr::select(USUBJID, ARM)
  ae_arm <- ae |> dplyr::inner_join(dm_arm, by = "USUBJID")
  
  tab <- Tplyr::tplyr_table(ae_arm, ARM) |>
    Tplyr::set_pop_data(dm) |>
    Tplyr::set_pop_treat_var(ARM) |>
    Tplyr::add_layer(
      Tplyr::group_count("All subjects") |>
        Tplyr::set_distinct_by(USUBJID) |>
        Tplyr::set_format_strings(Tplyr::f_str("xx", distinct_total))
    ) |>
    Tplyr::add_layer(
      Tplyr::group_count("Subjects with adverse events") |>
        Tplyr::set_distinct_by(USUBJID) |>
        Tplyr::set_format_strings(Tplyr::f_str("xx (xx %)", distinct_n, distinct_pct))
    ) |>
    Tplyr::add_layer(
      Tplyr::group_count(AESEV, by = "Adverse event severity") |>
        Tplyr::set_distinct_by(USUBJID) |>
        Tplyr::set_format_strings(Tplyr::f_str("xx (xx %)", distinct_n, distinct_pct))
    ) |>
    Tplyr::add_layer(
      Tplyr::group_count("Subjects with serious AE", where = AESER == "Y") |>
        Tplyr::set_distinct_by(USUBJID) |>
        Tplyr::set_format_strings(Tplyr::f_str("xx (xx %)", distinct_n, distinct_pct))
    )
  return(tab)
}

# create secod build function 
build_func2 <- function(tab) {
  Tplyr::build(tab, metadata = TRUE) |>
    Tplyr::apply_row_masks(row_breaks = TRUE)
}

# put second table into the output_list
output_list <- list(
  "Demographic" = list(
    tplyr_tab_fun = my_tplyr_fun,
    build_fun = build_func
  ),
  "Adverse Events" = list(
    tplyr_tab_fun = my_tplyr_fun2,
    build_fun = build_func
  )
)

# put everything together 
module_list <- list(
  "Table" = dv.tables::mod_Tplyr_table(
    module_id = "tplyr_table",
    output_list = output_list
  )
)
# start app
dv.manager::run_app(
  data = list("My data" = list(ae = ae, dm = dm)),
  module_list = module_list,
  filter_data = "dm",
  enable_dataset_filter = TRUE
)
  
```

As a last step you can also add a standalone listing to the module

```{r, eval=FALSE}
# Add additional entry for standalone listings
output_list <- list(
  "Demographic" = list(
    tplyr_tab_fun = my_tplyr_fun,
    build_fun = build_func
  ),
  "Adverse Events" = list(
    tplyr_tab_fun = my_tplyr_fun2,
    build_fun = build_func
  ),
  "Listing" = list(
    dataset_names = c("dm", "ae")
  )
)
# put everything together 
module_list <- list(
  "Table" = dv.tables::mod_Tplyr_table(
    module_id = "tplyr_table",
    output_list = output_list
  )
)
# start app
dv.manager::run_app(
  data = list("My data" = list(ae = ae, dm = dm)),
  module_list = module_list,
  filter_data = "dm",
  enable_dataset_filter = TRUE
)
```

